### 2025-02-27 notes/sec/set-indexed-product.tex
Simpler notation for \prod of a family.

\subsection{Set-indexed products}

$C$ \emph{has set-indexed products} if, for any set $I$ and any $I$-indexed family $X$ of objects of $C$,
there exists an object $\prod_{i \in I}X_i$ (also written $\prod_{I} X$) and family of morphisms $\eval_i$ in
$C$ such that for any object $Y$ and any family of morphisms $\{f_i: Y \to X_i\}_{i \in I}$ in $C$, there
exists a unique morphism $\lambda f$ making the following diagram commute:

\begin{center}
\begin{tikzcd}
   \prod_{I} X \arrow[r, "\eval_i"] & X_i \\
   Y \arrow[ru, "f_i"'] \arrow[u, dotted, "\lambda(f)"]
\end{tikzcd}
\end{center}

Note that $f$ is a morphism in $\Fam(I,C)$ from the constant family $\{Y\}_{i \in I}$ to $X$.