### 2025-03-08 notes/sec/biproduct.tex
More general statement of prod/coprod lemma.

\subsection{Biproducts}

\begin{definition}[Zero object]
A \emph{zero} object is an object which is both terminal and initial.
\end{definition}

Suppose $C$ a $\CMon$-enriched category.

\begin{definition}[Biproduct]
For any objects $A, B$, the \emph{biproduct} of $A$ and $B$ is an object $A \biprod B$ of $C$ together with
morphisms

\begin{center}
\begin{tikzcd}
   A \arrow[r, "\inj_A", shift left] &
   A \biprod B \arrow[l, "\proj_A", shift left] \arrow[r, "\proj_B"', shift right] &
   B \arrow[l, "\inj_B"', shift right]
\end{tikzcd}
\end{center}

\noindent satisfying

\begin{minipage}[t]{0.45\textwidth}
\begin{center}
\begin{salign*}
   \proj_A \comp \inj_A &= \id_A \\
   \proj_B \comp \inj_A &= \zero_{A,B}
\end{salign*}
\end{center}
\end{minipage}%
\begin{minipage}[t]{0.45\textwidth}
\begin{center}
\begin{salign*}
   \proj_B \comp \inj_B &= \id_B \\
   \proj_A \comp \inj_B &= \zero_{B,A}
\end{salign*}
\end{center}
\end{minipage}

\noindent and

\begin{salign*}
\inj_A \comp \proj_A + \inj_B \comp \proj_B &= \id_{A \biprod B}
\end{salign*}
\end{definition}

\begin{proposition}
\item
\begin{itemize}
\item If $C$ has biproducts then it has products given by $(A \biprod B, \proj_A, \proj_B)$ and coproducts
given by $(A \biprod B, \inj_A, \inj_B)$.
\item If $C$ has products (or coproducts) then they are biproducts.
\item If $C$ has a terminal (or initial) object then it is a zero object.
\end{itemize}
\end{proposition}

\noindent We say that $C$ is \emph{semi-additive} if it has finite biproducts.

\begin{proposition}
\label{prop:biproduct:prod-coprod}
Suppose $C$ a category with biproducts. Then the following diagram in $C$ commutes for any $f_1, g_1, f_2,
g_2$:

\begin{center}
\begin{tikzcd}[row sep=3em]
   & A' \\
   A \arrow[ur, "f_1"] \arrow[r, "\inj_A", shift left] &
   A \biprod B \arrow[l, "\proj_A", shift left] \arrow[r, "\proj_B"', shift right] \arrow[u, pos=0.3, "\coprodM{f_1}{g_1}"'] &
   B \arrow[l, "\inj_B"', shift right] \arrow[ul, "g_1"'] \\
   & B' \arrow[ul, "f_2"] \arrow[u, pos=0.6, "\prodM{f_2}{g_2}"'] \arrow[ur, "g_2"']
\end{tikzcd}
\end{center}
\end{proposition}