### 2025-03-14 notes.tex
secrefs.

\documentclass[acmsmall,nonacm]{acmart}

\usepackage{bbm}
\usepackage{tikz-cd}

\geometry{paperwidth=8.3in, paperheight=11.7in} % force to A4 for now
\settopmatter{printacmref=false}

\input{macros}

\begin{document}

\title{Approximation as Differentiation notes}
\maketitle

\section{Overview}

Notions covered so far:
\begin{itemize}
\item $\CMon$-enriched categories (\secref{cmon-enriched})
\item biproducts (\secref{biproduct})
\item category $\Fam(I,C)$ of $I$-indexed families of objects of $C$ (\secref{fam})
\item set-indexed products (\secref{set-indexed-product})
\item Grothendieck construction and category of families $\Fam(C)$ (\secref{grothendieck})
\item category of $\LatGal$ of bounded lattices and Galois connections (\secref{galois})
\end{itemize}

\noindent $\Set$ will usually be $\Setoid$ in the Agda implementation but we will gloss that detail for now.

\input{notes/sec/notation}

\section{Definitions}

\input{notes/sec/cmon-enriched}
\input{notes/sec/biproduct}
\input{notes/sec/fam}
\input{notes/sec/set-indexed-product}
\input{notes/sec/grothendieck}
\input{notes/sec/galois}

% \input{notes/language}

\section{Some references}

\cite{karvonen2020}

\bibliographystyle{plainnat}
\bibliography{bib}

\end{document}