### 2025-03-31 notes/sec/some-categories-with-biproducts.tex
Rename.

\subsection{Useful categories with biproducts}
\label{sec:categories-with-biproducts}

\subsubsection{Category of bounded lattices and Galois connections}
\label{sec:categories-with-biproducts:latgal}

Define the category $\LatGal$ where:
\begin{itemize}
\item objects $X, Y$ are the bounded lattices;
\item morphisms $f \vdash g: X \to Y$ are meet-semilattice homomorphisms $f: X \to Y$ paired with
join-semilattice homomorphisms $g: Y \to X$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in X$ and any
$y \in Y$.
\end{itemize}

\noindent $\LatGal$ has finite biproducts.

\subsubsection{Category of finite-dimensional vector spaces}
\label{sec:categories-with-biproducts:fdvect}

% \subsubsection{$\CMon$-valued presheaves over $\LatGal$}
%
% A presheaf $F: \LatGal^\op \to \CMon$ assigns to every bounded lattice $L$ a commutative monoid $F(L)$ and to
% every Galois connection $f: L \to L'$ a commutative monoid homomorphism $F(f): F(L) \to F(L')$.
% $\Func{\LatGal^\op}{\CMon}$ is the category of presheaves of commutative monoids on $\LatGal$.