### 2025-03-31 notes/sec/some-categories-with-biproducts.tex
Onto vector space over a field F.

\subsection{Useful categories with biproducts}
\label{sec:categories-with-biproducts}

\subsubsection{Category of bounded lattices and Galois connections}
\label{sec:categories-with-biproducts:latgal}

\begin{definition}[Bounded lattice]
A \emph{bounded lattice} $X = (X, \meet, \top, \join, \bot)$ is a partial order $X$ equipped with binary
operations $\meet$ and $\join$ and distinguished elements $\top$ and $\bot$ such that $(X, \meet, \top)$ is a
bounded meet-semilattice and $(X, \join, \bot)$ is a bounded join-semilattice.
\end{definition}

\noindent Since $(X, \meet, \top)$ is a commutative monoid, and thus a $\CMon$-enriched poset category, for
any $X, Y$ there is a zero meet-semilattice homomorphism $0_{X,Y}: X \to \top_Y$ and addition of homomorphisms
$f + g: X \to Y$ given by pointwise application of $\meet$. The dual construction applies to $(X, \join,
\bot)$. Now define the category $\LatGal$ where:
\begin{itemize}
\item objects $X, Y$ are the bounded lattices;
\item morphisms $f \dashv g: X \to Y$ are meet-semilattice homomorphisms $f: X \to Y$ paired with
join-semilattice homomorphisms $g: Y \to X$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in X$ and any
$y \in Y$.
\end{itemize}

\noindent and composition is component-wise.

$\LatGal$ is itself $\CMon$-enriched. For any $X, Y$ we have that $\LatGal(X,Y)$ is a commutative monoid,
where:
\begin{itemize}
\item $0_{X,Y}$ is the Galois connection $0_{X,Y} \dashv 0_{Y,X}: X \to Y$
\item $(f \dashv g) + (f' \dashv g') = (f + f') \dashv (g + g'): X \to Y$.
\end{itemize}

\noindent and biadditivity of composition is component-wise. Moreover $\LatGal$ has finite products, with the
trivial 1-point lattice as terminal object, product lattice $X \times Y$ as binary product, and product of
Galois connections defined component-wise; it follows that $\LatGal$ has finite biproducts
(\propref{biproduct:from-product-or-coproduct}).

\subsubsection{Category of finite-dimensional vector spaces}
\label{sec:categories-with-biproducts:fdvect}

\begin{definition}[Vector space over a field $F$]
Suppose $F$ a field with addition of elements written $a + b$ and multiplication written $a \mult b$.
Overloading $+$ and $\mult$, define a \emph{vector space $V = (V, +, \mult)$ over $F$} to be a set $V$
equipped with vector addition $+: V^2 \to V$ () and scalar multiplication $\mult: F \times V \to V$ where
$(V,+)$ is an Abelian group and the vector operations are compatible with the field operations in the
following sense:
\begin{itemize}
\item
\end{itemize}
\end{definition}

% \subsubsection{$\CMon$-valued presheaves over $\LatGal$}
%
% A presheaf $F: \LatGal^\op \to \CMon$ assigns to every bounded lattice $L$ a commutative monoid $F(L)$ and to
% every Galois connection $f: L \to L'$ a commutative monoid homomorphism $F(f): F(L) \to F(L')$.
% $\Func{\LatGal^\op}{\CMon}$ is the category of presheaves of commutative monoids on $\LatGal$.