### 2025-04-15 appendix/notes.tex
mathcal for metavariables that range over categories.

\section{Notes}

\subsection{Overview}

Covered so far:
\begin{itemize}
\item $\CMon$-enriched categories, including enriched functors and enriched Yoneda embedding (\secref{cmon-enriched})
\item biproducts and semi-additive categories (\secref{biproduct})
\item category $\Fam(I,\cat{C})$ of $I$-indexed families of objects of $\cat{C}$ (\secref{fam})
\item set-indexed products (\secref{set-indexed-product})
\item Grothendieck construction $\Grothendieck{\cat{C}}F$ and category of families $\Fam(\cat{C})$
(\secref{grothendieck})
\item Examples of useful semi-additive categories (\secref{useful-semi-additive-categories}):
   \begin{itemize}
   \item category $\LatGal$ of bounded lattices and Galois connections (\secref{categories-with-biproducts:latgal})
   \item category $\FinVect_F$ of finite-dimensional vector spaces over a field $F$
   (\secref{categories-with-biproducts:fdvect})
   \end{itemize}
\end{itemize}

\noindent $\Set$ will usually be $\Setoid$ in the Agda implementation but we will gloss that detail for now.

\input{appendix/notes/notation}
\input{appendix/notes/cmon-enriched}
\input{appendix/notes/biproduct}
\input{appendix/notes/fam}
\input{appendix/notes/set-indexed-product}
\input{appendix/notes/grothendieck}
\input{appendix/notes/useful-semi-additive-categories}

% \input{appendix/notes/language}