### 2025-04-15 appendix/notes/fam.tex

\subsection{Category of $I$-indexed families}
\label{sec:fam}

\begin{definition}[Category of $I$-indexed families of objects]
For any set $I$ and any category $\cat{C}$ write $\Fam(I,\cat{C})$ for the category where:
\begin{itemize}
\item objects are the $I$-indexed families of objects of $\cat{C}$;
\item morphisms from $X$ to $Y$ are families of morphisms $f_i: X_i \to Y_i$ in $\cat{C}$ for any $i \in I$.
\end{itemize}
\end{definition}

\noindent Equivalently $\Fam(I,\cat{C})$ is the functor category $\Func{I}{\cat{C}}$, interpreting $I$ as a
discrete category.

\begin{definition}[Constant family]
For any object $X$ of $\cat{C}$ write $\const_I(X)$ for the constant family $\{X\}_{i \in I}$.
\end{definition}

\begin{definition}[Reindexing]
For any $f: I \to J$ the \emph{reindexing} functor $\reindex{-}{f}: \Fam(J,\cat{C}) \to \Fam(I,\cat{C})$ sends
any $J$-indexed family $X$ to the $I$-indexed family $X[f]$ where $X[f]_i = X_{f(i)}$ for any $i \in I$, and
similarly for morphisms.
\end{definition}

\begin{definition}[$\Fam(-,\cat{C})$ functor]
$\Fam(-,\cat{C}): \Set^{\op} \to \Cat$ is then the functor which sends any set $I$ to $\Fam(I,\cat{C})$ and
any function $f: I \to J$ to the functor $\reindex{-}{f}: \Fam(J,\cat{C}) \to \Fam(I,\cat{C})$.
\end{definition}

\begin{proposition}
If $\cat{C}$ has limits (resp.~colimits) then $\Fam(I,\cat{C})$ has limits (resp.~colimits) computed
pointwise.
\end{proposition}