### 2025-04-15 appendix/notes/set-indexed-product.tex

\subsection{Set-indexed products}
\label{sec:set-indexed-product}

$\cat{C}$ \emph{has set-indexed products} if, for any set $I$ and any $I$-indexed family $X$ of objects of
$\cat{C}$, there exists an object $\prod_{i \in I}X_i$ (also written $\prod_{I} X$) and family of morphisms
$\eval_i$ in $\cat{C}$ such that for any object $Y$ and any family of morphisms $\{f_i: Y \to X_i\}_{i \in I}$
in $\cat{C}$, there exists a unique morphism $\lambda(f)$ making the following diagram commute for every $i
\in I$:

\begin{center}
\begin{tikzcd}
   \prod_{I} X \arrow[r, "\eval_i"] & X_i \\
   Y \arrow[ru, "f_i"'] \arrow[u, dotted, "\lambda(f)"]
\end{tikzcd}
\end{center}

Note that $f$ is a morphism $\const_{}(Y) \to X$ in $\Fam(I,\cat{C})$.