### 2025-04-16 appendix/notes/notation.tex
Standardise on X,Y,Z,W for objects.

\section{Notation}

In a category with coproducts or products we write:
\begin{itemize}
\item $\coprodM{f}{g}: X + Y \to Z$ for the coproduct of morphisms $f: X \to Z$ and $g: Y \to Z$
\item $\prodM{f}{g}: X \to Y \times Z$ for the product of morphisms $f: X \to Y$ and $g: X \to Z$
\item $(f,g): X \times Y \to Z \times W$ for $\prodM{f \comp \proj_1}{g \comp \proj_2}$ where $f: X \to Z$ and $g: Y \to W$.
\end{itemize}