### 2025-04-18 appendix/notes/preliminaries.tex
Preliminaries.

\section{Preliminaries}

\subsection{Notation}

\paragraph{Coproduct and product of morphisms.}

In a category with coproducts $X + Y$ or products $X \times Y$, we write:
\begin{itemize}
\item $\coprodM{f}{g}: X + Y \to Z$ for the coproduct of morphisms $f: X \to Z$ and $g: Y \to Z$
\item $\prodM{f}{g}: X \to Y \times Z$ for the product of morphisms $f: X \to Y$ and $g: X \to Z$
\item $(f,g): X \times Y \to Z \times W$ for $\prodM{f \comp \proj_1}{g \comp \proj_2}$ where $f: X \to Z$ and $g: Y \to W$.
\end{itemize}

\paragraph{Constant function.}

Write $\const(y): X \to Y$ for the function which sends any $x \in X$ to $y \in Y$.

\subsection{Enriched categories}