### 2025-07-02 main/higher-order.tex
Remove old higher-order section.

\section{Higher-Order Programs}
\label{sec:higher-order}