### 2025-07-03 notes/stable-coproducts.tex
Start on stable coproducts.

\subsection{Stable coproducts}
\label{sec:stable-coproducts}

Important to the idea of an extensive category is how coproducts behave under composition with an arbitrary
morphism. This is an instance of the notion of stability of colimits (in this case, coproducts) under
pullback or change of base.