### 2025-07-08 main/higher-order.tex
Reinstate missing semantics.

\section{Interpreting Higher-Order Programs}
\label{sec:higher-order}