### 2025-07-09 main.tex

\documentclass[acmsmall,screen,anonymous]{acmart}

\usepackage{tabularx}
\usepackage{bbm}
\usepackage{bbold}
\usepackage{mathpartir}
\usepackage{subcaption}
\usepackage{tikz-cd}
\usepackage{xspace}

%\geometry{paperwidth=8.3in, paperheight=11.7in} % force to A4 for now
\settopmatter{printacmref=false}
\citestyle{acmauthoryear}
\raggedbottom

\input{macros}

\begin{document}

\title{Approximation as Differentiation}

\author{Robert Atkey}
\email{robert.atkey@strath.ac.uk}
\orcid{0000-0002-4414-5047}
\affiliation{%
  \institution{University of Strathclyde}
  \city{Glasgow}
  \country{UK}}

\author{Roly Perera}
\email{roly.perera@cl.cam.ac.uk}
\orcid{0000-0001-9249-9862}
\affiliation{%
  \institution{University of Cambridge}
  \city{Cambridge}
  \country{UK}
}
\additionalaffiliation{%
   \institution{University of Bristol}
   \city{Bristol}
   \country{UK}
}

\begin{abstract}
\todo{abstract needed}
\end{abstract}
\maketitle

\input{main/introduction}
\input{main/approx-as-tangents}
\input{main/models}
\input{main/first-order}
\input{main/language}
\input{main/definability}
\input{main/related-work}
\input{main/conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{bib}

% \pagebreak
% \appendix
% \input{appendix/notes}

\end{document}