### 2025-07-10 main/definability.tex
Suggestion to give F a more meaningful name.

\section{Correctness of the Higher-Order Interpretation}
\label{sec:definability}

Introduction: what correctness properties we want; why we can't
compute the left adjoints from the forward maps (no completeness).

\begin{theorem}[\citet{fiore-simpson99}]
  \label{thm:glr-definability}
  Let $\cat{C}$ be an extensive bicartesian category, $\cat{D}$ be a
  bicartesian closed category, and $F : \cat{C} \to \cat{D}$ a functor
  preserving finite products and coproducts. Then there is a category
  $\GLR(F)$ and functors $p : \GLR(F) \to \cat{D}$ and
  $\hat{F} : \cat{C} \to \GLR(F)$, such that:
  \begin{enumerate}
  \item $\GLR(\cat{D},F)$ is bicartesian closed;
  \item $F = p \circ \hat{F} : \cat{C} \to \cat{D}$;
  \item The functor $p$ strictly preserves the bicartesian closed structure; and
  \item The functor $\hat{F}$ is full and preserves the bicartesian structure.
  \end{enumerate}
\end{theorem}

\begin{theorem}
  \label{thm:language-definability}
  For all $\Gamma \vdash M : \tau$, where $\Gamma$ and $\tau$ are
  first-order, there exists a morphism
  $g \in \Fam(\LatGal)(\sem{\Gamma}_{\mathit{fo}},
  \sem{\tau}_{\mathit{fo}})$ such that
  $\sem{\Gamma \vdash M : \tau} = F(g)$.
\end{theorem}

\roly{Give embedding $F$ a more specific name to clarify we're referring to \propref{ho-embedding} (esp.~as
\thmref{glr-definability} quantifies over arbitrary $F$)}

\begin{remark}
  We have formalised \thmref{glr-definability} and
  \thmref{language-definability} in Agda, except for the fact that
  $\Fam(C)$ is extensive\footnote{For reviewers: we intend to fix this
    before any final version.}. Our formalisation uses E-categories
  (\cite{palmgren-wilander}) so we cannot state
  \thmref{glr-definability} as given because the concept of strict
  preservation does not make sense. Therefore, the formalised version
  of \thmref{language-definability} is ``up to isomorphism''.

  We believe that this result also extends to the list type, since it
  is constructed from an infinite coproduct, but have not yet checked
  all the details.
\end{remark}

\begin{proof}
  Instantiate \thmref{glr-definability} with
  $F : \Fam(\LatGal) \to \Fam(\MeetSLat \times \JoinSLat^\op)$. By
  \propref{XXX} we know that $F$ preserves finite products and
  coproducts. $\Fam(\LatGal)$ has extensive coproducts because it is
  the free coproduct completion. Thus we get the category $\GLR(F)$ in
  which we can interpret the calculus with the property that this
\end{proof}

FIXME: another property we might want: The interpretations in
$\Fam(\MeetSLat)$ and $\Fam(\MeetSLat \times \JoinSLat^\op)$ agree at
first order.