### 2025-07-10 main/language.tex
Rough pass.

\section{Higher-Order Language}
\label{sec:language}

To model Galois slicing semantically for higher-order programs, we define a simple total functional
programming language over a signature of base types and primitive operations, extending the simply-typed
lambda calculus.

\subsection{Syntax}
\label{sec:language:syntax}

The language is parameterised by a signature $\Sigma = (\PrimTy, \PrimOp)$ consisting of a set $\PrimTy$ of
base types $\rho$ and a family of sets $\PrimOp^\rho_{\rho_1,\ldots,\rho_n}$ of primitive operations $\phi$ of
arity $n$ over those base types. The syntax is defined in \figref{syntax}. Types includes base types $\rho$
drawn from $\PrimTy$, along with standard type formers for sums, products, functions and lists, plus a type
constructor $\tyLift$ to allow approximation points to be added explicitly, as discussed in
\secref{models-of-total-gps}. Terms include variables, the usual introduction and elimination forms, a monadic
return and bind for lifted terms, and primitive operations $\phi$.

The language is intentionally minimal: it excludes general recursion, and general inductive or coinductive
types, which we will consider in future work (\secref{conclusion}). Typing judgments for terms are standard
and shown in \figref{typing}, with the usual rules for products, sums, functions, lists and lifting.

\input{fig/syntax}
\input{fig/typing}

\subsection{Semantics}
\label{sec:language:semantics}

\input{fig/semantics}

A model $M$ of $\Sigma$ can be given in any category $\cat{C}$ with finite products, and assigns to each base
type $\rho \in \PrimTy$ an object $\sem{\rho}_{\PrimOp}$ in $\cat{C}$, and to each primitive operation $\phi
\in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}_{\Op}: \sem{\rho_1}_{\PrimTy} \times \ldots
\times \sem{\rho_n}_{\PrimTy} \to \sem{\rho}_{\PrimTy}$. A model of $\tyList$ can also be given in such a
category, and consists of an endofunctor $\List: \cat{C} \to \cat{C}$, plus for any object $X$, morphisms
$\nil: 1 \to \List(X)$ and $\cons: X \times \List(X) \to \List(X)$, and for any objects $Y, Z$ and morphisms
$f_\nil: Z \to Y$ and $f_\cons: Z \times X \times Y \to Y$, a morphism $\fold(f_\nil,f_\cons): Z \times
\List(X) \to Y$, with $\nil$, $\cons$ and $\fold$ satisfying the usual laws for lists.

The semantics for the full language is then given in terms of an arbitrary bicartesian closed category $\Sem$,
models of $\Sigma$ and $\tyList$ in $\Sem$, and a strong monad $(\Lift, \eta, \mu)$ with strength
$\mathsf{st}$ for interpreting $\tyLift$. \figref{semantics:types} gives the interpretations of types $\tau$
and contexts $\Gamma$ as objects; \figref{semantics:terms} gives the interpretation of terms $\Gamma \vdash t:
\tau$ as morphisms $\sem{\Gamma} \to \sem{\tau}$.

\subsubsection{Interpretation for higher-order \GPS}

Given a model $M$ of $\Sigma$ in $\Fam(\LatGal)$, the category $\Fam(\MeetSLat \times \JoinSLat^\op)$ can
implement $\Sem$. It is bicartesian closed by \corref{mslat-jslat-bcc}, and can model $\tyList$ using infinite
coproducts (\secref{models-of-total-gps:coproducts-and-products}). The embedding $\HoEmbed: \Fam(\LatGal) \to
\Fam(\MeetSLat \times \JoinSLat^\op)$ from \propref{ho-embedding} preserves finite products, providing a model
$H(M)$ of $\Sigma$. Finally the lifting monad in $\Fam(\LatGal)$ (\secref{first-order:lifting}) is also
preserved by $H$. The correctness of this interpretation is considered in the next section.