### 2025-07-10 main/conclusion.tex
Some blurb on source transformation approach.

\section{Conclusion and Future Work}
\label{sec:conclusion}

\paragraph{Integrating XAI techniques with Galois slicing.} \roly{Sketch what this might look like.}
Explainable AI (XAI) techniques like Gradient-weighted Class Activation Mapping (Grad-CAM)~\cite{selvaraju20}
use reverse-mode AD selectively on specific activations, calculating \emph{heat maps} that highlight input
regions contributing to a given classification decision. It might be possible to compose these techniques with
\GPS whilst still preserving useful round-tripping properties; if so this could be a basis for end-to-end
explainability in systems that combine neural networks with traditional symbolic computation.

\paragraph{More Approximation Domains.} The intervals example is
interesting, but is it useful, and are there more examples like this?
One route might be to follow \cite{edalat-heckmann98} and explore
whether metric spaces (which already provide a native notion of
approximation) can be embedded in $\Fam(\LatGal)$.

\paragraph{Alternative Higher-order Models}

1. $\Fam(\PSh_{\CMon}(\LatGal))$

2. Concrete sheaves over $\Fam(\LatGal)$ analogous to Diffeological spaces?

\paragraph{Categorical Models of Differentation}

\paragraph{Recursion and Partiality.}
1. Despite using stable domain theory as our starting point, we have ignored

2. Conjecture that the category with $(X : \namedcat{Cpo}, \partial X : X \to \CLatGal)$, where $\CLatGal$ is the category of complete lattices and galois connections is the right setting. Would need to investigate the connections with dI-domains, bidomains, and bistable bidomains (\cite{laird07}).

\paragraph{Implementation}

% To formalise this, we'd need to... See also $\Fam(\PSh_{\CMon}(\namedcat{Syn}))$?

An interesting implementation technique, as an alternative the denotational approach presented here or the
trace-based approaches used in earlier \GPS implementations, would be to develop a source-to-source
transformation, in direct analogy with the CHAD approach to automatic differentiation
\cite{vakar22,nunes2023}. In their approach, forward and reverse-mode AD are implemented as compositional
transformations on source code, guided by a universal property: they arise as the unique structure-preserving
functors from the source language to a suitably structured target language formalised as a Grothendieck
construction. Adapting this to Galois slicing would allow slicing to ``compiled in'', avoiding the need for a
custom interpreter and presenting opportunities for optimisation.