### 2025-07-11 main/approx-as-tangents.tex
[ tex ] further space saving

\section{Approximations as Tangents}
\label{sec:approx-as-tangents}

We motivate our approach by showing how to combine ideas from differential geometry and stable domain theory to reconstruct the ideas of \GPS in a denotational setting.
% From this basis, we apply the CHAD framework of V{\'a}k{\'a}r \etal to obtain a denotational model of automatic differentiation that gives us an effective way to interpret programs as functions along with their forward and backwards approximation maps.

\subsection{Manifolds, Smooth Functions, and Automatic Differentiation}
\label{sec:approx-as-tangents:autodiff}

% For what follows we do not need to know in detail what a manifold is, or exactly how smooth functions are defined. We only state some definitions in order to fix terminology and to justify our claim of a connection to manifolds and smooth functions. \bob{Need a good reference here.}

The general study of differentiable functions takes place on
\emph{manifolds}, topological spaces that ``locally'' behave like an
open subset of the Euclidean space $\RR^n$. The spaces $\RR^n$
themselves are manifolds, but so are ``non-flat'' examples such as
$n$-spheres and yet more exotic spaces. Every point $x$ in a manifold
$M$ has an associated \emph{tangent vector space} $\tangents_x(M)$
consisting of linear approximations of curves on the manifold passing
through $x$. Each point also has a \emph{cotangent vector space}
$\cotangents_x(M) = \tangents_x(M) \linearto \RR$. The tangent and
cotangent spaces are finite dimensional, so in the presence of a
chosen basis they are canonically isomorphic. In the case when the
manifold is $\RR^n$, then every tangent space is isomorphic to $\RR^n$
as well.

Smooth functions $f$ between manifolds $M$ and $N$ are functions on their points that are locally differentiable on $\RR^n$. Manifolds and smooth functions form a category $\Man$. Each smooth function induces maps of the (co)tangent spaces:
\begin{itemize}[leftmargin=\enummargin]
\item The \emph{forward derivative} (tangent map, pushforward) $\pushf{f}_x$ is a linear map $\tangents_x(M) \linearto \tangents_{f(x)}(N)$. In the Euclidean case when $M = \RR^m$ and $N = \RR^n$, the tangent map can be represented by the Jacobian matrix of partial derivatives of $f$ at $x$.
\item The \emph{backward derivative} (cotangent map, pullback) $\pullf{f}_x$ is a linear map $\cotangents_{f(x)}(N) \linearto \cotangents_x(M)$. In the Euclidean case, the backward derivative is represented by the transpose of the Jacobian of $f$ at $x$.
\end{itemize}

\begin{remark}[Chain Rule]
  \label{rem:chain-rule}
  A useful property of derivative maps is that they compose according
  to the chain rule. Suppose that $f : M \to N$ and $g : N \to K$ are
  smooth functions. Then for any $x \in M$, we have:
  \begin{itemize}
  \item $\pushf{(g \circ f)}_x = \pushf{g}_{f(x)} \circ \pushf{f}_x : \tangents_x(M) \linearto \tangents_{g(f(x))}(K)$
  \item $\pullf{(g \circ f)}_x = \pullf{f}_x \circ \pullf{g}_{f(x)} : \cotangents_{g(f(x))}(K) \linearto \cotangents_x(M)$
  \end{itemize}
  The chain rule has the practical effect that we can compute
  derivative maps of $f$ and $g$ independently and compose them,
  instead of the potentially more difficult task of computing the
  derivative maps of $g \circ f$. As we shall see below, stable maps
  also obey a chain rule, and this forms the basis of the general
  categorical approach to differentiability that we describe in
  \secref{models-of-total-gps}.
\end{remark}

Computing the forward and backward derivatives of smooth functions $f$ has many applications of practical interest. For example, computation of the reverse derivative is of central interest in machine learning by gradient descent, the main technique used to train deep neural networks~\cite{rumelhart88,goodfellow16}.

Derivatives can be computed numerically by computing $f$ on small perturbations of its input, or symbolically by examining a closed-form representation of $f$. However, a more common and practical technique is to use \emph{automatic differentiation}, where a program computing $f$ is instrumented to produce (a representation of) the forward and/or backward derivative as a side-effect of producing the output~\cite{linnainmaa76}. This has led to the area of differentiable programming, where programming languages and their implementations are specifically designed to admit efficient automatic differentation algorithms~\cite{jax2018github,abadi16,elliott17,sigal24}.

\subsection{Stable Functions as Differentiable Functions}

Our thesis is that \GPS is a generalised form of differentiable programming, where tangents are not linear approximations of curves but instead are qualitative information approximations of elements. Smooth functions in this setting are Berry's \emph{stable functions} \cite{berry79,berry82}. We now introduce these concepts and how they relate to \GPS.

\subsubsection{Domains as a Qualitative Theory of Approximation}

Domain theory is a method for defining the semantics of programs that
handle infinite data such as functions or infinite streams. Domains
are certain partially ordered sets where the ordering denotes a
relationship of qualitative information content: if $x \sqsubseteq y$,
then $y$ may contain more information than $x$. For example, if $x$
and $y$ are functions, then $y$ may be defined at more values than
$x$. Infinite objects are understood in terms of their approximations
in this sense, and domains are assumed to be closed under least upper
bounds (lubs) of directed sets, meaning that any internally consistent
collection of elements has a ``completion'' that contains all the
information covered by the set. Programs are interpreted as monotone
functions that preserve directed lubs. Monotonicity captures the idea
that if the input gets more defined, then the output can get more
defined. Preservation of lubs, or \emph{continuity}, states that a
function interpreting a program cannot act inconsistently on
approximations and their completion, which corresponds to the
intuitive idea that a function that is computable cannot look at a
non-finite amount of input to produce a finite output. \citet{abramsky-jung} provide a comprehensive introduction to domain theory.

For the purposes of \GPS, we are interested in using
approximations not to model computation on infinite objects, but instead
for revealing how programs explore their inputs when producing parts
of their output. Therefore, we ignore completeness properties of the
partially ordered sets we consider.

\subsubsection{Bounded Meets and Conditional Multiplicativity}
\label{sec:bounded-meets-and-cm}

When giving a denotational semantics for sequential programming
languages, Scott-continuous functions are too permissive. Famously,
\citet{plotkin77lcf}'s Parallel OR (\exref{parallel-or}, below) is
continuous but does not explore its input in a way consistent with a
sequential implementation. Continuous functions can explore their
input in a non-deterministic way as long as the result is
deterministic. This non-determinism results in functions whose
output cannot be assigned a unique minimal input that accounts for
it. Therefore, continuous functions are incompatible with the central
idea in \GPS that we should be able to identify a \emph{minimal} part
of the input that leads to a part of the output.

\emph{Stability} is a property that can be required of monotone functions, that was invented by \citet{berry79} in an attempt to capture sequentiality. This was unsuccessful (see the $\mathrm{gustave}$ function in \exref{parallel-or}), but we will see now how it is closely related to the problem of computing forward and backwards maps of approximations needed in \GPS. A textbook description of stable functions in the context of domain theory is given by \citet[Chapter 12]{amadio-curien}. We start with a property that is weaker than stability, but is easier to motivate in connection with derivatives of smooth functions.

In a partially ordered set $X$, for any $x \in X$ the set of elements below $x$, $\downset{x} = \{ x' \mid x' \sqsubseteq x \}$, is itself a partially ordered set. These approximations of $x$ we will think of as ``tangents'' at $x$, and the whole set $\downset{x}$ as the ``tangent space''. Tangent spaces are vector spaces, so in the partially ordered setting we take elements of $\downset{x}$ to be approximations of processes defined at $x$. As we can add tangents, we assume we can take meets of approximations in $\downset{x}$:

% \footnote{\bob{Is it possible to derive the tangent spaces by
% thinking about continuous functions passing through $x$?}}

\begin{definition}
  A \emph{bounded meet poset} is a partially ordered set $X$ where for
  every $x \in X$, $\downset{x}$ is a meet semilattice, with $x$ as
  the top element.
\end{definition}

For the approximation version of the forward derivative of $f$ at $x$, we take $f$'s restriction to $\downset{x}$, taking approximations of the input to approximations of the output. Matching the linearity of the forward derivative of smooth functions, we require that these restrictions preserve meets. This is exactly the definition of \emph{conditionally multiplicative} function from \citet{berry79}:

\begin{definition}
  A \emph{conditionally multiplicative} (cm) function $f : X \to Y$ is
  a monotone function such that for all $x \in X$, the restriction
  $f_x : \downset{x} \to \downset{f(x)}$ preserves meets.
\end{definition}

\begin{theorem}
  Bounded meet lattices and conditionally multiplicative functions form
  a category $\CM$. This category has products, coproducts, and
  exponentials.
\end{theorem}

\begin{proof}
  \AGDA. See \citet[Theorem 12.1.9]{amadio-curien} for the
  case when the posets are also cpos. The crucial technical step,
  identified by \citet{berry79}, is that the ordering on conditionally
  multiplicative functions is not the extensional ordering
  ($f \sqsubseteq_{\mathrm{ext}} g$ iff
  $\forall x. f(x) \sqsubseteq g(x)$) but instead the stable ordering:
  $f \sqsubseteq_{\mathrm{st}} g$ iff $f \sqsubseteq_{\mathrm{ext}} g$ and
  $\forall x, x'.\,x \sqsubseteq x' \Rightarrow f(x) = f(x') \land
  g(x)$.
\end{proof}

\begin{remark}[Chain Rule]
  \label{rem:chain-rule-cm}
  It is almost a triviality at this point, but the crucial point is
  that, for any cm functions $f : X \to Y$ and $g : Y \to Z$, the
  restriction maps (``forward derivatives'') compose according to the
  chain rule from \remref{chain-rule}:
  \begin{displaymath}
    (g \circ f)_x = g_{f(x)} \circ f_x
  \end{displaymath}
  We will see this phenomenon repeated in the definition of stable
  functions below.
\end{remark}

\begin{example}[Conditionally Multiplicative Functions]
  \label{ex:strict-short-circuit}
  To see the effect of conditional multiplicativity, consider several
  ways of defining the OR on the lifted booleans
  $\mathbb{B}_\bot$. Two functions that are cm are the strict and
  left-short-circuiting ORs\footnote{The clauses in these examples are
    shorthand for the graph of the function. They are not to be
    understood as pattern matching clauses in a language like Haskell,
    where it is not possible to match on $\bot$.}:
  \begin{displaymath}
    \begin{array}[t]{l@{(}l@{,~}l@{)~}c@{~}l}
      \mathrm{strictOr}&\mathsf{tt}&\mathsf{tt}&=&\mathsf{tt} \\
      \mathrm{strictOr}&\mathsf{tt}&\mathsf{ff}&=&\mathsf{tt} \\
      \mathrm{strictOr}&\mathsf{ff}&\mathsf{tt}&=&\mathsf{tt} \\
      \mathrm{strictOr}&\mathsf{ff}&\mathsf{ff}&=&\mathsf{ff} \\
      \mathrm{strictOr}&\bot&\_&=&\bot \\
      \mathrm{strictOr}&\_&\bot&=&\bot \\
    \end{array}
    \qquad
    \begin{array}[t]{l@{(}l@{,~}l@{)~}c@{~}l}
      \mathrm{shortCircuitOR}&\mathsf{tt}&\_&=&\mathsf{tt} \\
      \mathsf{shortCircuitOR}&\mathsf{ff}&x&=&x \\
      \mathsf{shortCircuitOR}&\bot&\_& =& \bot
    \end{array}
    \qquad
    \raisebox{-15ex}{
      \begin{tikzpicture}
        \node (top) at (0,0) {$(\mathsf{tt},\mathsf{ff})$};
        \node [below of=top,xshift=-0.5cm] (oi) {$(\bot,\mathsf{ff})$};
        \node [below of=top,xshift=0.5cm] (io) {$(\mathsf{tt},\bot)$};
        \node [below of=oi,xshift=0.5cm] (oo) {$(\bot,\bot)$};
        \draw (top) -- (oi);
        \draw (top) -- (io);
        \draw (oi) -- (oo);
        \draw (io) -- (oo);
      \end{tikzpicture}
    }
  \end{displaymath}
  In the poset $\mathbb{B}_\bot^2$, a typical poset of approximations of a fully defined element is shown to
  the right. For $\mathrm{strictOr}$, any approximation that isn't the fully defined input is mapped to
  $\bot$, while $\mathrm{shortCircuitOr}$ maps the partially defined $(\mathsf{tt},\bot)$ to $\mathsf{tt}$.
  Thus, even though these functions operate identically on fully defined inputs, they differ in their
  \emph{derivatives} on partially defined input, exposing how they explore their arguments differently. That
  they are cm can be checked by examining their restrictions' behaviour. If we take the approximations
  $(\bot,\mathsf{ff})$ and $(\mathsf{tt},\bot)$, then their meet is $(\bot,\bot)$; $\mathrm{strictOr}$ maps
  all three elements to $\bot$, so is cm here since $\bot \wedge \bot = \bot$; and $\mathrm{shortCircuitOR}$
  has $(\bot,\mathsf{ff}) \mapsto \bot$ and $(\mathsf{tt},\bot) \mapsto \mathsf{tt}$, the meet of which is
  $\bot = \mathrm{shortCircuitOR}(\bot,\bot)$. Other combinations can be checked similarly.
\end{example}

\begin{example}[A non-Conditionally Multiplicative Function]
  \label{ex:parallel-or}
  A function that is not cm is Plotkin's Parallel OR \cite{lcf77}, which short-circuits in both arguments. It returns $\mathsf{tt}$ if either argment is $\mathsf{tt}$ even if the other argument is not defined:
  \begin{displaymath}
    \begin{array}{l@{(}l@{,~}l@{)~}c@{~}l}
      \mathrm{parallelOR}&\mathsf{tt}&\_&=&\mathsf{tt} \\
      \mathrm{parallelOR}&\_&\mathsf{tt}&=&\mathsf{tt} \\
      \mathsf{parallelOR}&\mathsf{ff}&\mathsf{ff}&=&\mathsf{ff} \\
      \mathsf{parallelOR}&\bot&\bot&=&\bot
    \end{array}
  \end{displaymath}
  We have $\mathrm{parallelOR}(\mathsf{tt}, \bot) \wedge \mathrm{parallelOR}(\bot, \mathsf{tt}) = \mathsf{tt} \wedge \mathsf{tt} = \mathsf{tt}$ but $\mathrm{parallelOR}((\mathsf{tt},\bot) \wedge (\bot, \mathsf{tt})) = \mathrm{parallelOR}(\bot, \bot) = \bot$, so it is not cm.

  Parallel OR is famous because it is not \emph{sequential}, meaning intuitively that it cannot be implemented without running the two arguments in parallel to see if one of them returns $\mathsf{tt}$. The fact that it exists in the standard domain theoretic semantics of PCF means that this semantics is incomplete for reasoning about observational equivalence in PCF. Since Parallel OR is not cm, one might hope that cm-ness is enough to capture sequentiality, and hence potentially give a fully abstract model of PCF. However, the following ternary function $\mathbb{B}_\bot^3 \to \{\top,\bot\}$ is cm but admits no sequential implementation that fixes an order that the arguments are examined in:
  \begin{displaymath}
    \begin{array}{l@{(}l@{,~}l@{,~}l@{)~}c@{~}l}
      \mathrm{gustave}&\mathsf{tt}&\mathsf{ff}&\_&=&\top \\
      \mathrm{gustave}&\mathsf{ff}&\_&\mathsf{tt}&=&\top \\
      \mathrm{gustave}&\_&\mathsf{tt}&\mathsf{ff}&=&\top \\
      \mathrm{gustave}&\_&\_&\_&=&\bot
    \end{array}
  \end{displaymath}
  Due to the way that the cases are defined, there is no way of constructing a pair of approximations for which preservation of their meet does not hold. In terms of derivatives, this makes sense in that we are only concerned about the intensional behaviour of a function at a point and its approximations. Parallel OR has two incompatible approximation behaviours at the point $(\mathsf{tt},\mathsf{tt})$. The $\mathrm{gustave}$ function does have consistent behaviour at each approximation for each point.
\end{example}

\begin{example}[Intervals and Maximal Elements]
  \label{ex:intervals-and-maxima-elements}
  The set of intervals
  $\{ [l,u] \in \mathbb{R} \times \mathbb{R} \mid l \leq u \}$ ordered
  by reverse inclusion forms a (Scott) domain
  \cite{scott-outline}. The set of maximal elements is exactly
  $\mathbb{R}$. This domain has been proposed as a model of
  approximate real number computation \cite{escardo-real-pcf}. The
  information approximation reading is intuitive: as intervals move up
  the order they become tighter, containing more information about the
  number they are approximating.

  Given this reading, it makes sense to wonder if we can use interval
  approximations as ``information tangents'' of real numbers, where
  derivatives take approximating intervals to approximating
  intervals. Since intervals form a Scott domain, they are bounded
  complete and hence have bounded meets. However, the addition
  function on intervals, $[l_1,u_1] + [l_2,u_2] = [l_1+l_2,u_1+u_2]$,
  is not conditionally multiplicative, as can be easily checked.

  A solution is to use the set of intervals with nominated points that
  they are approximating: $\{[l,x,u] \mid l \leq x \leq u\}$. The
  ordering now is that $[l_1,x_1,u_1] \sqsubseteq [l_2,x_2,u_2]$ iff
  $x_1 = x_2$ and $l_1 \leq l_2$ and $u_2 \leq u_1$. Consequently, the
  maximal elements are $[x,x,x]$, recovering $\mathbb{R}$ again, but
  the approximations of each number all form independent sub-lattices.
  Addition is defined as
  $[l_1,x_1,u_1] + [l_2,x_2,u_2] = [(l_1+x_2) \sqcap
  (l_2+x_1),x_1+x_2,(u_1+x_2)\sqcup(u_2+x_1)]$, which is conditionally
  multiplicative. Note how this definition bears a resemblance to the
  product rule for derivatives, with (in the lower end of the
  interval) $\sqcap$ replacing $+$.

  This example is important to us because it shows that (for total
  programs) we are separately interested in the maximal elements and
  their approximations, and that approximations of each maximal
  element may have to be considered separately.

  Relatedly, \citet{edalat-heckmann98} proposed a Scott domain of
  formal balls on a metric space, where again the maximal elements are
  the points of the original space. More generally,
  \citet[Section V-6]{continuous-lattice-book} describe \emph{domain
    environments}, which are domains whose maximal elements are
  exactly the points of a topological space. We are not aware of any
  work linking domain environments in general to stable functions. It
  would be interesting to see whether the situation for intervals,
  where approximations must be relative to a nominated point, repeats
  in general when considering conditionally multiplicative functions.
\end{example}

Given these examples, conditional multiplicativity seems to be a
reasonable analogue to functions with a well-defined notion of
derivative. For \GPS we also require an analogue to the reverse
derivative, where we map approximations backwards to give the least
approximation of the input for a given approximation of the output. In
the case of smooth functions, we are always guaranteed a reverse
derivative. However, there is not always a best way to map
approximations backwards for cm functions, as the following example
shows.

\begin{example}[Is Conditional Multiplicativity Enough?]
  \label{ex:non-stable-function}
  An example that is conditionally multiplicative, but does not admit
  a backwards map of approximations (from \citet[just before Lemma
  12.2.3]{amadio-curien}, originally due to Berry) is given by
  $\mathrm{unstable} : D \to \{\bot \sqsubseteq \top\}$, where
  $D = \bot \sqsubseteq \cdots \sqsubseteq n \sqsubseteq \cdots
  \sqsubseteq 1 \sqsubseteq 0$, as $\mathrm{unstable}(\bot) = \bot$
  and $\mathrm{unstable}(n) = \top$. This is monotone, and preserves
  meets in every $\downset{x}$. But there is no ``best'' (i.e., least)
  input that gives us any finite output.
\end{example}

\subsubsection{Stable functions and L-posets}
\label{sec:stab}

In light of the \exref{non-stable-function}, we turn to
\citet{berry79}'s definition of \emph{stable function} that requires
the existence of a reverse mapping directly, even without assuming
that any meets exist:

\begin{definition}[Stable function]
  Let $f : X \to Y$ be a monotone function between posets $X$ and
  $Y$. The function $f$ is \emph{stable} if for all $x \in X$ and
  $y \leq f(x)$:
  \begin{enumerate}
  \item (\textsc{Existence}) there exists an $x_0 \leq x$ such that $y \leq f(x_0)$, and
  \item (\textsc{Minimality}) for any $x'_0 \leq x$ such that $y \leq f(x'_0)$ then
    $x_0 \leq x'_0$ .
  \end{enumerate}
\end{definition}

\begin{example}\label{ex:stable-functions}
  \leavevmode\\[-3ex]
  \begin{enumerate}[leftmargin=\enummargin]
  \item The function $\mathrm{strictOr}$ is stable. For example, for
    the input-output pair
    $(\mathsf{tt},\mathsf{ff}) \mapsto \mathsf{tt}$, the minimal input
    that gives this output is exactly $(\mathsf{tt}, \mathsf{ff})$. If
    we take the approximation $\bot \leq \mathsf{tt}$ of the output,
    then the corresponding minimal input is $(\bot, \bot)$. The
    function $\mathrm{shortCircuitOR}$ is also stable. For the
    input-output pair $(\mathsf{tt},\mathsf{ff}) \mapsto \mathsf{tt}$,
    the minimal input that gives this input is $(\mathsf{tt},\bot)$,
    indicating that the presence of $\mathsf{ff}$ in the second
    argument was not necessary to produce this output. As with
    $\mathrm{strictOr}$, the minimal input required to produce the
    output $\bot \leq \mathsf{tt}$ is again $(\bot,\bot)$.

  \item The $\mathrm{parallelOR}$ function is not stable. For the
    input-output pair $(\mathsf{tt},\mathsf{tt}) \mapsto \mathsf{tt}$,
    there is no one minimal input that produces this output. We have
    both $\mathrm{parallelOR}(\mathsf{tt},\bot) = \mathsf{tt}$ and
    $\mathrm{parallelOR}(\bot,\mathsf{tt}) = \mathsf{tt}$, which are
    incomparable and their greatest lower bound $(\bot,\bot)$ gives
    the output $\bot$.

  \item The $\mathrm{gustave}$ function is stable. Despite there being
    no one minimal input that achieves the output $\top$, each of the
    minimal inputs that can achieve this output are pairwise
    incomparable, so for each specific input that gets output $\top$
    there is a unique minimal input that achieves it (listed in the
    first three lines of the definition).

    In terms of \GPS the $\mathrm{gustave}$ function does not present
    a problem. For any particular run (i.e., input $\mapsto$ output
    pair), there is an unambiguous minimal input that achieves the
    output, no matter that it was not achieved by a sequential
    processing of the input.

  \item As discussed above, $\mathrm{unstable}$ is not stable, but is
    conditionally multiplicative.

  \item The addition function on intervals with nominated points in
    \exref{intervals-and-maxima-elements} is stable. Given the input
    $[l_1,x_1,u_1], [l_2,x_2,u_2]$ and an approximation
    $[l,x_1+x_2,u]$ of the output, the minimal approximations of the
    input are $[l-x_2,x_1,u-x_2], [l-x_1,x_2,u-x_1]$. We can read this
    as saying if the output was the maximal element $x_1+x_2$ but we
    only require the output to be in the range $[l,x_1+x_2,u]$, then
    we can obtain intervals containing the input values that are
    enough to obtain the desired output approximation \emph{assuming
      that the other input is kept the same}. Note the analogy to
    partial derivatives in multi-variable calculus, where the
    derivative is computed in each variable independently.
  \end{enumerate}
\end{example}

% The fact that these two functions' stability witnesses reveal
% information about how they depend on their input is what we will
% exploit in order to use the idea of stability for \GPS. It is
% interesting that sequentiality is not necessarily required in order
% to obtain backwards maps of approximations, as might be assumed by
% thinking about how to trace computations for provenance information.

Stability has an alternative definition in terms of Galois
connections, which will be more useful for what follows. This
characterisation is due to~\citet{taylor99}. We first define Galois
connections, which we used informally in \exref{introduction-example}.

\begin{definition}[Galois connection]
  \label{def:galois-connection}
  Suppose $X$ and $Y$ are posets. A \emph{Galois connection} $f \adj g: X \to Y$ is a pair of monotone functions $f: Y \to X$ and $g: X \to Y$ satisfying $y \leq g(x) \iff f(y) \leq x$ for any $x \in X$ and $y \in Y$. Since a Galois connection is also an adjunction, we refer to $f$ as the left adjoint and $g$ as the right adjoint.
\end{definition}

\begin{lemma}
  A monotone function $f : X \to Y$ is stable if and only if for all
  $x \in X$, the restriction of $f_x : \downset{x} \to \downset{f(x)}$
  has a left Galois adjoint.
\end{lemma}

\begin{proof}
  If $f$ is stable, then define a left adjoint
  $f^*_x : \downset{f(x)} \to \downset{x}$ by setting $f^*_x(y)$ to be
  the minimal $x_0$ required by stability. This is monotone: if
  $y \leq y'$, then we know that $y \leq y' \leq f(f^*_x(y'))$ by the
  definition of $f^*_x$, so $f^*_x(y) \leq f^*_x(y')$ by minimality of
  $f^*_x(y)$. For the adjointness, let $x' \leq x$ and $y \leq
  f(x)$. Then if $f^*_x(y) \leq x'$, we have
  $y \leq f(f^*_x(y)) \leq f(x')$ by monotonicity of $f$ and the first
  part of stability. In the other direction, if we have
  $y \leq f(x')$, then by uniqueness we have $f^*_x(y) \leq x'$.

  If, for every $x$, $f_x$ has a left adjoint $f^*_x$, then for any
  $x', y$ we have $y \leq f_x(x') \Leftrightarrow f^*_x(y) \leq
  x'$. So $f^*_x(y)$ is the element that satisfies
  $y \leq f(f^*_x(y))$, and it is minimal since if $y \leq f_x(x'_0)$
  then $f^*_x(y) \leq x'_0$.
\end{proof}

Even though stable functions can be defined on any partially ordered
set, in light of the analogy with tangent spaces it makes sense to
require that meets, preserved by forward approximation maps, and
joins, preserved by backwards approximation maps, exist:

\begin{definition}
  An \emph{L-poset} is a partially ordered set $X$ such that for every $x \in X$, the principal downset $\downset{x}$ is a bounded lattice (i.e., have all finite meets and joins).
\end{definition}

This lemma is an instance of standard facts about Galois connections
preserving meets and joins:

\begin{lemma}
  For L-posets $X$ and $Y$, a stable function $f : X \to Y$ preserves
  meets in its forward part $f_x$ and joins in its reverse part
  $f^*_x$.
\end{lemma}

The converse to this lemma (that functions that preserve meets in their forward part have a left Galois adjoint) is not true, as was demonstrated by the non-stable function in \exref{non-stable-function}. In the case when the posets $\downset{x}$ are \emph{complete}, and $f_x$ preserves infinitary meets, then we are guaranteed a left Galois adjoint. In that example, the infinite set $\{0, 1, 2, \dots, n, n-1, \dots\}$ not including $\bot$ of approximations of $0$ does not have a greatest lower bound, so the order is not complete.

\begin{theorem}
  L-posets and stable functions form a category $\Lposet$ with products
  and coproducts.
\end{theorem}

\begin{remark}[Chain Rule]
  \label{rem:chain-rule-stable}
  As for the ``forward derivatives'' of conditionally multiplicative
  functions, the forward and backwards parts of a stable function
  compose according to the chain rule (c.f. \remref{chain-rule}):
  \begin{itemize}
  \item $(g \circ f)_x = g_{f(x)} \circ f_x : \downset{x} \linearto \downset{g(f(x))}$
  \item $(g \circ f)^*_x = f*_x \circ g^*_{f(x)} : \downset{g(f(x))} \linearto \downset{x}$
  \end{itemize}
  We will use this property in \propref{embed-stable} to show that
  $\Lposet$ embeds into our category of sets-with-approximation.
\end{remark}

The category of L-posets and stable functions is not cartesian
closed. To make it so, we would need to require that the principal
downsets $\downset{x}$ are \emph{complete} lattices. \citet[Theorem
12.5.10]{amadio-curien} details the proof. Intuitively, to generate
the best approximation of an input value for a function, we need to
take the infimum over all possible input values.

Our goal is to model a higher-order language suitable for writing
queries on databases as in \exref{introduction-example}, so why should
we not just take complete L-posets and stable functions as our model
of \GPS?  We have two reasons for moving to a different model in
\secref{models-of-total-gps}:

\begin{enumerate}[leftmargin=*]
\item Even without completeness, in bounded meet posets and L-posets
  values and their approximations live in the same set. However, in
  the \emph{total} query language we wish to model in
  \secref{language}, we are not directly interested in the behaviour
  of programs on approximations as we would be for partial programs
  with general recursion. (Moreover, in the light of
  \exref{intervals-and-maxima-elements} it is not clear whether
  approximations for partiality and approximations for stability ought
  to be the same thing. We discuss this further in
  \secref{conclusion}.) In \exref{intervals-and-maxima-elements},
  maximal elements could be taken to be the ``proper values''. One
  idea is to restrict to conditionally multiplicative or stable
  functions that preserve maximal elements. However, this idea fails
  at higher order: functions that take maximal elements to maximal
  elements are not themselves maximal elements. We could devise a
  category of L-posets with totality predicates (which would pick out
  maximal elements at first-order) and totality preserving functions,
  but we prefer a more direct method of separating values proper from
  their approximations using the \emph{Category of Families}
  construction as we explain in \secref{models-of-total-gps:fam}.
\item A more practical reason is that we wish to formalise our
  construction in the proof assistant Agda \cite{agda} in order to get
  an \emph{executable} model of the language in
  \secref{language}. Agda's type theory is both predicative and
  constructive. Predicativity means that it does not have complete
  lattices in the classical sense: for a type $X$ we can only get
  suprema and infima of families in a lower universe level than
  $X$. This is not necessarily a problem, as \citet{dejong21} show how
  to develop a large amount of domain theory in a predicative
  setting. However, due to constructivity the lifting construction
  (due to \citet{escardo-knapp}) requires a proof of definedness to
  extract a value. Since we are modelling a total language, we prefer
  to have a model that computes directly.
\end{enumerate}

\subsection{Summary}
\label{sec:diff-stab-summary}

We have seen that \citet{berry79}'s theory of stable functions between suitable partial orders can be seen as form of differentiability with forward and reverse derivatives. In the next section, we describe a model based on these ideas suitable for constructing executable models of total higher order languages.

We end this section with a conjecture. Although we have argued above
that there is an analogy between stable functions and smooth
functions, we have not stated any mathematical theorems substantiating
this. \emph{Tangent Categories} \cite{cockett14,cockett18} are a
categorical axiomatisation of the properties of manifolds and smooth
functions in terms of the presence of tangent bundles $T(X)$ for every
object $X$, forward derivatives and additivity of tangents. They
generalise Cartesian Differential Categories \cite{cdcs}, which are an
axiomatisation of Euclidean spaces and smooth functions.

\begin{conjecture}
  \label{con:tangent-stable-fns}
  (1) Bounded meet posets and conditionally multiplicative functions
  form a Tangent category where the tangent bundle
  $T(X) = \{(x,x') \mid x' \leq x\}$ and addition of tangents is given
  by meets. (2) L-posets and stable functions form a \emph{reverse
    Tangent category} \cite{reverse-tangents}.
\end{conjecture}

As well as codifying exactly what we mean by differentiable structure
on partially ordered sets, proving this conjecture would also tell us
what higher derivatives mean in this context as well, something that
we have not considered above. We will extend this conjecture to our
refined model of lattice approximated sets in \secref{models:summary}.

% Let us summarise our analogy between differentiable functions between manifolds and stable functions between L-posets:

% \begin{itemize}
% \item For \GPS, we assume that every value has an associated lattice of {\em approximations}. For differentiable programs, every point has an associated vector space of {\em tangents}.
% \item For \GPS, every program has an associated forward approximation map that takes approximations forward from the input to the output. This map {\em preserves meets}. For differentiable programs, every program has a forward derivative that takes tangents of the input to tangents of the output. The forward derivative map is {\em linear}, so it preserves addition of tangents and the zero tangent.
% \item For \GPS, every program has an associated backward approximation map that takes approximations of the output back to least approximations of the input. This map {\em preserves joins}. For differentiable programs, every program has a reverse derivative that takes tangents of the output to tangents of the input. This map is again {\em linear}.
% \item For \GPS, the forward and backward approximation maps are related by being a Galois connection. For differentiable programming, the forward and reverse derivatives are related by being each others' transpose.
% \end{itemize}